---
title: "Helper Functions"
author: "Michael Koohafkan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Helper Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This document gets illustrates some of the helper functions in `cimir `.

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE
)
```

First, simply load the `cimir` library:

```{r eval = FALSE}
library(cimir)
```

In this vignette, we'll use some example data from the 
Markleeville station (#246). The station metadata can be 
retrieved with `cimis_station()`:

```{r eval = FALSE}
station.meta = cimis_station(246)
print(station.meta)
```

```{r echo = FALSE, results = "markup"}
station.meta = cimis_station(246)
knitr::kable(station.meta)
```

Notice that the station latitude and longitude is provided as a text
string, in both Hour Minute Second (HMMS) and Decimal Degree (DD) 
format. We can extract one or the other of these formats using 
`cimis_format_location()`:

```{r eval = FALSE}
station.meta = cimis_format_location(station.meta, "DD")
head(station.meta)
```

```{r echo = FALSE, results = "markup"}
station.meta = cimis_format_location(station.meta, "DD")
knitr::kable(head(station.meta))
```

Now let's retrieve some data with `cimis_data()`:


```{r eval = FALSE}
station.data = cimis_data(246, "2017-04-01", "2017-04-30",
  c("day-air-tmp-avg", "hly-air-tmp"))
head(station.data)
```

```{r echo = FALSE, results = "markup"}
station.data = cimis_data(246, "2017-04-01", "2017-04-30",
  c("day-air-tmp-avg", "hly-air-tmp"))
knitr::kable(head(station.data))
```

Notice that hourly data returns timestamps in two columns "Date" and 
"Hour". Furthermore, since we requested both a daily item and an hourly 
item, the daily item records have `NA` values for the "Hour" column. We 
can collapse these columns into a single datetime column using 
`cimis_to_datetime()`:

```{r eval = FALSE}
station.data = cimis_to_datetime(station.data)
head(station.data)
```

```{r echo = FALSE, results = "markup"}
station.data = cimis_to_datetime(station.data)
knitr::kable(head(mutate(station.data, Datetime = paste(Datetime))))
```

Note that a time of `00:00:00` is used for daily records.

The CIMIS Web API has fairly conservative limitations on the number
of records you can query at once. Large queries can be split 
automatically into a series of smaller queries using `cimis_split_queries`:

```{r}
queries = cimis_split_query(247, "2017-04-01", "2018-04-30",
  c("day-air-tmp-avg", "hly-air-tmp"))
queries
```

The queries can then be run in sequence using e.g. `mapply()` or 
`purrr::pmap()`:

```{r eval = FALSE}
purrr::pmap_dfr(queries, cimis_data)
```

Note that the CIMIS API may reject your requests if you submit too many 
queries in a short period of time.
