---
title: "Quick Start"
author: "Michael Koohafkan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This document gets you up and running with `cimir `, an R interface to
the 
[California Irrigiation Management Information System](https://cimis.water.ca.gov/)
(CIMIS) [Web API](http://et.water.ca.gov/Home/Index). In order to use 
this package, you will need to 
[create a CIMIS account](https://cimis.water.ca.gov/Auth/Register.aspx) 
and [request a web services AppKey](http://et.water.ca.gov/Home/Register/).

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE
)
```

First, simply load the `cimir` library:

```{r eval = FALSE}
library(cimir)
```

If you want `cimir` to automatically use an existing AppKey when you attach
the library, you can either set the option `cimir.appkey` in your 
`.Rprofile` or create an environment variable `CIMIS_APPKEY`. 
To set an AppKey for a single session, use `set_key()`:

```{r eval = FALSE}
set_key("YOUR-APP-KEY-HERE")
```

There are two types of CIMIS queries: station queries and data queries.
Station queries return metadata on stations, while data queries return
environmental monitoring data. For a list of possible data items, use 
the function  `cimis_items()`:

```{r}
cimis_items()
```

To get a list of station IDs, use  `cimis_station()`:

```{r}
cimis_station()
```

To get metadata for a subset of stations, simply pass a vector of 
station IDs. You can also query metadata for stations associated 
with a particular zipcode (or spatial zipcode) 
`cimis_zipcode()` and `cimis_spatial_zipcode()`:

```{r}
cimis_zipcode()
cimis_spatial_zipcode()
```

Once the stations or regions of interest are identified, 
you can retrieve data using `cimis_data()`:

```{r}
cimis_data(targets = 170, start.date = "2018-09-25", end.date = "2018-10-01")
```

The data are returned in tidy format, i.e. the `"Value"` column 
displays the recorded value for the data item specified in the 
`"Item"` column. The CIMIS Web API returns a predefined set of data 
items by default, but custom item sets can also be passed to 
`cimis_data()`. Data queries also support zipcodes, spatial coordinates 
(latitude and longitude, in decimal degrees) and street addresses, 
as well as additional options for data units and data providers. See 
`?cimis_data` and the 
[CIMIS Web API documentation](http://et.water.ca.gov/Rest/Index) for
more information. 

When hourly data is retrieved, an additional column `"Hour"` is 
returned. If both hourly and daily data items are requested,
the `"Hour"` column will be filled with `NA` values for the daily 
data items. The `cimis_to_datetime()` function provides a quick way to 
combine the columns `"Date"` and `"Hour"` to a single `"Datetime"` column. 
Note that when mixed daily and hourly data is present in the dataframe, 
the daily values are assumed to correspond to midnight (00:00).

```{r}
cimis_to_datetime(cimis_data(targets = 170, start.date = "2018-09-25",
  end.date = "2018-09-27", items = c("hly-soil-tmp", "day-air-tmp-avg")))
```

CIMIS data sometimes includes quality control flags in the `Qc` column that
indicate potential issues with returned data. Descriptions of these flags
can be retrieved `cimis_flags()`:

```{r}
cimis_flags()
```

Note that quality control flag designations were changed in 1993. For more
information, see the 
[CIMIS Data Overview - Quality Control](https://cimis.water.ca.gov/Resources.aspx).
web page.
